---
title: "covid 19"
author: "Natalia Lekach"
date: "3/17/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
# Introduction
## Question 1 - Is dying independent of sex?

Using COVID-19 data collected from January 1, 2020 to May 1, 2020 in South Korea, we will determine if the males die at a different rate than females or if death from COVID-19 is independent of sex. 

Null hypothesis: Dying from COVID-19 is independent of sex. Males and females die at the same rate.
Alternative hypothesis: Dying from COVID-19 is not independent of sex. Males and females die at a different rate. 
 
Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(result, sex))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this
```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```
  
Of the patients who survived from COVID-19, the majority (~57%) were females. Of the patients who died from COVID-19, ~20% more of them were males. Males consisted of just of 60% of the deaths and  females consisted of slightly less than 40% of the deaths.  

## Numerical results


## Inferential results 




## Conclusions

