---
title: "covid 19"
author: "Natalia Lekach"
date: "3/17/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
# Introduction
## Question 1 - Is dying independent of sex?

Using COVID-19 data collected from January 1, 2020 to May 1, 2020 in South Korea, we will determine if the males die at a different rate than females or if death from COVID-19 is independent of sex. 

Null hypothesis: Dying from COVID-19 in South Korea is independent of sex. Males and females die at the same rate.
Alternative hypothesis: Dying from COVID-19 in South Korea is not independent of sex. There is a relationships between deaths in South Korea from COVID-19 and sex. 
 
Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well.  
  
Table of patients:
```{r}
kRdata$Sex = as.factor(kRdata$Sex)
kRdata$result = as.factor(kRdata$result)
# cross tabulation, patient state by gender

xtabs(~Sex+result,data=kRdata)
```
  
# Descriptive Results  
  
## Graphical Summaries
  
##### Using ggplot commands, we created two bar charts - one based on frequency and the other on percent.
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
Of the patients who survived from COVID-19, the majority (~57%) were females. Of the patients who died from COVID-19, ~20% more of them were males. Males consisted of above 60% of the deaths and females consisted of slightly less than 40% of the deaths.  

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
Of the patients who survived from COVID-19, the majority (56%) were females. Of the patients who died from COVID-19, 18% more of them were males. Males consisted of above 60% of the deaths and females consisted of less than 40% of the deaths. 

# Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

With a relatively small sample size for those who died from COVID-19 in this dataset, we will examine the Fisher Exact Test scores. With a small p-value of 0.0009, we can determine that there is a statistically significant relationship between sex and mortality from COVID-19. The odds ratio indicates that males are 2.3 more likely to die from COVID-19 than females. 

# Conclusions


With a low p-value, we must determine that there is not enough evidence to support the null hypothesis. It is unlikely that these results occured by chance. This means that there was a statistically significant difference in the mortality of males and the mortality of females from COVID-19 in South Korea from January 1. 2020 to May 1, 2020. In fact, males were 2.3 times more likely to die from COVID-19 than females.   